---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = FALSE
)
```

# TestGenerator

<!-- badges: start -->
<!-- badges: end -->

Creates tests on patient data for DARWIN EU studies. It accepts an Excel file with a sample of patients, and tests on an empty CDM if the results are correct from a particular study package.

## Installation

You can install the development version of TestGenerator from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("darwin-eu-dev/TestGenerator")
```

## Example

Function to read the .xls file with data generally from 10 patients. By default, it generates a JSON files in the inst/testCases folders. 

```{r example}
library(TestGenerator)

TestGenerator::readPatients(here::here("extras", "RSV_Test_Data.xlsx"))

```

Data then can be pushed to a blank CDM with the complete vocabulary.

```{r cars}

cdm <- TestGenerator::patientCDM()

```
