---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = FALSE
)
```

# TestGenerator

<!-- badges: start -->
<!-- badges: end -->

Does my cohort is including the correct patients? Am I calculating an intersection in the right way? Is that the expected value for treatment duration? It just takes one incorrect parameter to get incoherent results in a pharmacoepidemiological study, and it is challenging to test calculations on huge and complex databases. 

That is why TestGenerator is useful to push a micro sample of around 10 patients to unit test a study on the OMOP-CDM. It includes tools to create a blank CDM with a complete vocabulary and check if the code is doing what we expect. 

This package is based on the unit testing of the [Eramus MC Ranitidine Study](https://github.com/mi-erasmusmc/RanitidineStudy/blob/master/unitTesting_README.md).

## Installation

You can install the development version of TestGenerator from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("darwin-eu-dev/TestGenerator")
```

## Example

The user should create an Excel file with a micro population of around 10 patients for testing purposes. That can include any table from the OMOP-CDM. 

`readPatients()` will read the Excel file, and saves the data in a JSON file. This is useful if the user wants to create more than one sets of test populations for testing. 

```{r, eval=FALSE}

TestGenerator::readPatients(
  filePath = "~/pathto/test_data.xlsx",
  sampleName = "test",
  tables = c("person", "observation_period", "drug_exposure", "condition_occurrence",
    "visit_occurrence", "visit_context", "visit_detail", "death"),
  outputPath = "inst/testCases"
)

```

`patientCDM()` pushes one of those test populations into a blank CDM reference. 

```{r}

cdm <- TestGenerator::patientCDM(
  pathJson = "inst/testCases", 
  sampleName = "test")

```

Now the user has a CDM reference with a complete vocabulary and a universe of just 10 patients to unit test functions of a particular study. 
