---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(dplyr)
library(testthat)
```

# TestGenerator

<!-- badges: start -->
[![R-CMD-check](https://github.com/darwin-eu/TestGenerator/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/darwin-eu-dev/TestGenerator/actions/workflows/R-CMD-check.yaml)
[![codecov](https://codecov.io/github/darwin-eu/TestGenerator/branch/main/graph/badge.svg)](https://app.codecov.io/github/darwin-eu-dev/TestGenerator?branch=main)
[![CRAN status](https://www.r-pkg.org/badges/version/TestGenerator)](https://CRAN.R-project.org/package=TestGenerator)
<!-- badges: end -->

Does my cohort picked the correct number patients? Am I calculating an intersection in the right way? Is that the expected value for treatment duration? It just takes one incorrect parameter to get incoherent results in a pharmacoepidemiological study, and it is very challenging to test calculations on huge and complex databases. 

That is why TestGenerator is useful to push a small sample of patients to unit test a study on the OMOP-CDM. It includes tools to create a blank CDM with a complete vocabulary and check if the code is doing what we expect in very specific cases.  

This package is based on the unit testing written for the [Eramus MC Ranitidine Study](https://github.com/mi-erasmusmc/RanitidineStudy/blob/master/unitTesting_README.md).

## Installation

To install TestGenerator:

```{r, eval=FALSE}
# CRAN version

install.packages("TestGenerator")
```

## Example

The user can provide an Excel file [(link to sample)](https://github.com/darwin-eu-dev/TestGenerator/raw/main/inst/extdata/icu_sample_population.xlsx) or a set of CSV files that represent tables of the OMOP-CDM, with a micro population of just 8 patients for testing purposes. 

`readPatients()` will read either Excel or CSVs, and then saves the data in a JSON file. This is useful if the user wants to create more than one Unit Test Definitions. If the parameter `outputPath` is `NULL` The files are saved in the `testthat/testCases` folder of the package. Alterna

```{r, eval=FALSE}
TestGenerator::readPatients(filePath = "~/pathto/testPatients.xlsx",
                            testName = "test",
                            outputPath = NULL,
                            cdmVersion = "5.3")
```

Alternatively, the user can use the functions `readPatients.xl` or `readPatients.csv` directly. 

```{r, eval=FALSE}
TestGenerator::readPatients.xl(filePath = "~/pathto/testPatients.xlsx",
                               testName = "test",
                               outputPath = NULL,
                               cdmVersion = "5.3")

TestGenerator::readPatients.csv(filePath = "~/pathto/csv/files",
                                testName = "test",
                                outputPath = NULL,
                                cdmVersion = "5.3",
                                reduceLargeIds = FALSE)
```

`patientCDM()` pushes one of those Unit Test Definitions into a blank CDM reference with a complete version of the vocabulary. If the `pathJSON` parameter is `NULL`, `TestGenerator` will look for the JSON test files in the `testthat/testCases` folder. 

```{r, eval=FALSE}
cdm <- TestGenerator::patientsCDM(pathJson = NULL, 
                                  testName = "test",
                                  cdmVersion = "5.3")

```

Now the user has a CDM reference with a complete vocabulary and just 8 patients.

```{r}
filePath <- system.file("extdata/icu_sample_population.xlsx", 
                        package = "TestGenerator")

outputPath <- file.path(tempdir(), "test")

dir.create(outputPath)

TestGenerator::readPatients(filePath = filePath, 
                            testName = "test", 
                            outputPath = outputPath, 
                            cdmVersion = "5.3")
cdm <- TestGenerator::patientsCDM(pathJson = outputPath, 
                                  testName = "test",
                                  cdmVersion = "5.3")

cdm$person
```

The reference can be used to create a cohort and create unit tests. 

```{r}
test_cohorts <- system.file("extdata",
                            "test_cohorts",
                            package = "TestGenerator")

cohort_set <- CDMConnector::readCohortSet(test_cohorts)
  
cdm <- CDMConnector::generate_cohort_set(cdm,
                                         cohort_set,
                                         name = "test_cohorts")

cohortAttrition <- CDMConnector::attrition(cdm[["test_cohorts"]])
  
excluded_records <- cohortAttrition %>% 
    pull(excluded_records) %>% 
    sum()
  
expect_equal(excluded_records, 0)
```

With `graphCohort()` it is possible to visualise the timeline for particular patient. 

```{r}

diazepam <- cdm[["test_cohorts"]] %>% 
  filter(cohort_definition_id == 1) %>% 
  collect()

icu_visit <- cdm[["test_cohorts"]] %>% 
  filter(cohort_definition_id == 2) %>% 
  collect()

TestGenerator::graphCohort(subject_id = 4, list("diazepam" = diazepam,
                                                "icu_visit" = icu_visit))

```

```{r, echo=FALSE}
unlink(outputPath, recursive = TRUE)
duckdb::duckdb_shutdown(duckdb::duckdb())
```
