---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# TestGenerator

<!-- badges: start -->
[![R-CMD-check](https://app.codecov.io/gh/darwin-eu-dev/TestGenerator/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/darwin-eu-dev/TestGenerator/actions/workflows/R-CMD-check.yaml)
[![codecov](https://app.codecov.io/gh/darwin-eu-dev/TestGenerator/branch/main/graph/badge.svg)](https://app.codecov.io/gh/darwin-eu-dev/TestGenerator)
<!-- badges: end -->

Does my cohort picked the correct number patients? Am I calculating an intersection in the right way? Is that the expected value for treatment duration? It just takes one incorrect parameter to get incoherent results in a pharmacoepidemiological study, and it is challenging to test calculations on huge and complex databases. 

That is why TestGenerator is useful to push a micro sample of around 10 patients to unit test a study on the OMOP-CDM. It includes tools to create a blank CDM with a complete vocabulary and check if the code is doing what we expect. 

This package is based on the unit testing written for the [Eramus MC Ranitidine Study](https://github.com/mi-erasmusmc/RanitidineStudy/blob/master/unitTesting_README.md).

## Installation

To install the development version of TestGenerator:

```{r, eval=FALSE}
# From CRAN
install.packages("TestGenerator")

# Development version
remotes::install_github("darwin-eu-dev/TestGenerator")
```

## Example

The user should provide an Excel file with a micro population of around 10 patients for testing purposes. That can include any table from the OMOP-CDM. 

`readPatients()` will read the Excel file, and saves the data in a JSON file. This is useful if the user wants to create more than one Unit Test Definitions. 

```{r, eval=FALSE}
TestGenerator::readPatients(
  filePath = "~/pathto/testPatients.xlsx",
  testName = "test",
  outputPath = "inst/testCases"
)

```

`patientCDM()` pushes one of those Unit Test Definitions into a blank CDM reference with a complete version of the vocabulary.

```{r, eval=FALSE}
cdm <- TestGenerator::patientCDM(
  pathJson = "inst/testCases", 
  testName = "test")

```

Now the user has a CDM reference with a complete vocabulary and a universe of just 10 patients to unit test functions of a particular study. 
