---
title: "Create Sample Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Create Sample Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE}
library(TestGenerator)
```

`TestGenerator` takes as an input an Excel file with sheets that represent a table in the OMOP-CDM. The following example [(testPatientsRSV.xlsx)](https://github.com/darwin-eu-dev/TestGenerator/raw/main/inst/extdata/testPatientsRSV.xlsx) represents a population of 10 patients, some of them with RSV.

```{r, echo=FALSE}
personTable <- readxl::read_xlsx(system.file("extdata/testPatientsRSV.xlsx", 
                                               package = "TestGenerator"),
                                               sheet = "person")
head(personTable, n = 10)
```
The user can include only the tables that are relevant to the analysis. 

```{r, echo=FALSE}
patientTables <- readxl::excel_sheets(system.file("extdata/testPatientsRSV.xlsx", 
                                                  package = "TestGenerator"))
patientTables
```

`TestGenerator::readPatients()` converts the file into JSON format and saves it in the project. The sample data is then pushed to a blank CDM with `patientsCDM()`. 

```{r, echo=FALSE}
filePath <- system.file("extdata/testPatientsRSV.xlsx", package = "TestGenerator")
outputPath <- file.path(tempdir(), "test")
dir.create(outputPath)
TestGenerator::readPatients.xl(filePath = filePath, outputPath = outputPath)
cdm <- TestGenerator::patientsCDM(pathJson = outputPath, testName = "test")
print(cdm)
unlink(outputPath, recursive = TRUE)
```

That returns a CDM reference object that now can be used to perform unit tests.

```{r, echo=FALSE}
print(cdm$person)
duckdb::duckdb_shutdown(duckdb::duckdb())
```

