---
title: "Remote databases and GitHub Actions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Remote databases and GitHub Actions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

DuckDB is the default backend in TestGenerator, and it is usually the fastest
place to start. Remote database testing is useful when you also want to check
that your study code works on another SQL engine. This is especially helpful
for code that will run in PostgreSQL, SQL Server, or Databricks/Spark in
production.

The idea is the same as with local testing: you create a small, controlled test
population, load it into a CDM, run your study code, and assert the expected
result. The difference is that `patientsCDM()` uploads the test CDM to a remote
database and returns a CDM reference connected to that backend.

## Use `patientsCDM()` with a remote database

First create a Unit Test Definition JSON file with `readPatients()` or
`readPatients.csv()`. Then call `patientsCDM()` with the backend you want to
test.

```{r postgresql-example}
library(TestGenerator)

cdm <- patientsCDM(
  pathJson = "tests/testthat/testCases",
  testName = "my_test_population",
  cdmVersion = "5.4",
  dbms = "postgresql"
)
```

The supported remote values for `dbms` are:

| Backend | `dbms` value |
| --- | --- |
| PostgreSQL | `"postgresql"` |
| SQL Server | `"sqlserver"` |
| Databricks/Spark | `"databricks"` |

For remote databases, `patientsCDM()` creates the CDM locally first, trims the
vocabulary to what the test population needs, uploads the result to the remote
database, and returns the remote CDM reference. By default, TestGenerator creates
a temporary test schema for the upload. You can also provide `writeSchema` if
you want to choose the schema name yourself.

```{r write-schema}
cdm <- patientsCDM(
  pathJson = "tests/testthat/testCases",
  testName = "my_test_population",
  cdmVersion = "5.4",
  dbms = "sqlserver",
  writeSchema = "testgenerator_my_case"
)
```

When the test has finished, clean up the remote schema and close the connection:

```{r cleanup-remote}
cleanupTestCdm(cdm)
```

This is important for shared remote databases. The test schemas are small, but
cleaning them up keeps the database tidy and avoids name conflicts in later
runs.

## Environment variables for direct use

When you call `patientsCDM()` directly against a remote database, TestGenerator
reads the connection details from environment variables.

| Backend | Required environment variables |
| --- | --- |
| PostgreSQL | `POSTGRESQL_SERVER`, `POSTGRESQL_DBNAME`, `POSTGRESQL_PORT`, `POSTGRESQL_USER`, `POSTGRESQL_PASSWORD` |
| SQL Server | `SQLSERVER_SERVER`, `SQLSERVER_DBNAME`, `SQLSERVER_PORT`, `SQLSERVER_USER`, `SQLSERVER_PASSWORD` |
| Databricks/Spark | `DATABRICKS_HOST`, `DATABRICKS_TOKEN`, `DATABRICKS_HTTPPATH` |

For Databricks, TestGenerator also reads `DATABRICKS_USER` and
`DATABRICKS_WORKSPACE` when they are set. If they are not set, it uses `token`
as the user and `hive_metastore` as the workspace/catalog.

For SQL Server, TestGenerator reads `SQL_SERVER_DRIVER` when it is set. If it is
not set, it uses `ODBC Driver 18 for SQL Server`.

A typical local `.Renviron` setup could look like this:

```{bash envvars, eval=FALSE}
POSTGRESQL_SERVER=localhost
POSTGRESQL_DBNAME=cdm
POSTGRESQL_PORT=5432
POSTGRESQL_USER=postgres
POSTGRESQL_PASSWORD=your-password

SQLSERVER_SERVER=localhost
SQLSERVER_DBNAME=cdm
SQLSERVER_PORT=1433
SQLSERVER_USER=sa
SQLSERVER_PASSWORD=your-password

DATABRICKS_HOST=https://your-workspace.cloud.databricks.com
DATABRICKS_TOKEN=your-token
DATABRICKS_HTTPPATH=/sql/1.0/warehouses/your-warehouse
```

Use values that match your own database or Databricks workspace. The database
user needs permission to create schemas, create tables, insert data, read data,
and drop the test schema during cleanup.

## Automate backend tests in GitHub Actions

The same remote database workflow can be used in continuous integration. The
goal is to let GitHub Actions prove that your package tests work on the database
engines you care about.

TestGenerator includes workflow templates for PostgreSQL, SQL Server, and
Databricks. You can copy all of them into your package with:

```{r all-actions}
useGithubAction(overwrite = TRUE)
```

You can also create only the workflow you need:

```{r one-action}
useGithubAction(dbms_type = "postgresql", overwrite = TRUE)
useGithubAction(dbms_type = "sqlserver", overwrite = TRUE)
useGithubAction(dbms_type = "databricks", overwrite = TRUE)
```

This writes workflow files to `.github/workflows`. Commit those files to your
repository so GitHub Actions can run them.

## What each action expects

The workflow templates run fixed test files. Your package must provide the file
for each backend you enable:

| Backend | Required test file |
| --- | --- |
| PostgreSQL | `tests/testthat/test-postgresql.R` |
| SQL Server | `tests/testthat/test-sqlserver.R` |
| Databricks/Spark | `tests/testthat/test-databricks.R` |

Each test file should call your package code, create the test CDM with
`patientsCDM(dbms = ...)`, run the study logic, and use `testthat` expectations
to check the result. If a test has an error or a failed expectation, the action
is intended to fail.

For PostgreSQL and SQL Server, the GitHub Actions workflow creates a Docker
container inside the action run. That means users do not need to configure
database secrets for these two workflows. The action sets the connection
environment variables itself and tests against the temporary container.

For Databricks, the action connects to your own remote Databricks workspace.
You must configure the required repository secrets in GitHub:

| GitHub secret | Purpose |
| --- | --- |
| `DATABRICKS_HOST` | URL of the Databricks workspace |
| `DATABRICKS_TOKEN` | Token used to connect |
| `DATABRICKS_HTTPPATH` | SQL warehouse or cluster HTTP path |

Optional Databricks secrets are:

| GitHub secret | Default when not set |
| --- | --- |
| `DATABRICKS_WORKSPACE` | `hive_metastore` |
| `DATABRICKS_ODBC_DRIVER_URL` | The default Simba Spark ODBC driver URL used by the workflow |

Because Databricks uses a real remote workspace, make sure the token has enough
permission to create schemas and tables, insert the test data, query the result,
and remove the test schema afterwards.

## A minimal backend test pattern

A backend-specific test usually follows this shape:

```{r test-pattern}
testthat::test_that("study logic works on PostgreSQL", {
  cdm <- NULL
  on.exit({
    if (!is.null(cdm)) {
      TestGenerator::cleanupTestCdm(cdm)
    }
  }, add = TRUE)

  cdm <- TestGenerator::patientsCDM(
    pathJson = "tests/testthat/testCases",
    testName = "my_test_population",
    cdmVersion = "5.4",
    dbms = "postgresql"
  )

  result <- myPackage::runMyStudy(cdm)

  testthat::expect_equal(result$n_subjects, 3)
})
```

The exact expectations should be specific to your micro population. Good tests
usually check counts, dates, cohort entry and exit, exclusions, or any other
result that should be predictable from the small input dataset.
